---
title: "Lab 9: Before Table Formatting"
author: "Daltzwer Oliveros"
format:
  html: 
    self-contained: true
    code-tools: true 
    code-fold: true 
execute:
  echo: true
  messages: false
  error: true
  eval: true
  warning: false
editor: visual
---

```{r set up}
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
library(gt)
library(DT)
baby_names <- read_csv(here("Week 9", "Labs", "StateNames_A.csv"))
```

# 3.1 Summarizing and Visualizing the Number of Allisons

```{r}
baby_names <- baby_names |>
  rename("Sex" = Gender)
```

```{r}
baby_names |>
  filter(Name == "Allison") |>
  group_by(State, Sex) |>
  summarise(Count = sum(Count), .groups = "keep") |>
   pivot_wider(names_from = Sex,
              values_from = Count,
              values_fill = 0) 
 
  
```

# 3.2 Creating Allison_f dataset

```{r}
allison_f <- baby_names |>
  filter(Sex == "F", Name == "Allison")
allison_f
```

# 3.3 Plot of the number of Babies named Allison 

```{r}
allison_f_year <- allison_f |>
  group_by(Year) |>
  summarise(Count = sum(Count)) 
ggplot(data = allison_f_year, mapping = aes(x = Year, 
                                            y = Count)
                                            ) +
  geom_line() +
  labs(x = "Year Observed", 
       y = " ",
       title = "Number of Female Born Babies Named Allison")

```

# 4.1 Modeling the Number of Allisons

```{r}
Allison_lm <- lm(Count ~ Year, data = allison_f_year)
Allison_lm
```

# 4.5 Regression Equation

```{r}
broom::tidy(Allison_lm)
```

Predicted number of Allisons = 209689.76 - 101.52 (year)

# 4.6 Residuals vs Fitted Plot

```{r}
Allison_lm |>
  broom::augment() |>
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```

I believe we do see some evidence of patterns in the graph above but not enough to cause concern since they are not the fanning and funneling shape we are looking for. 

# 4.7 Conclusion

Based on the information given to us above, we can see that for every year that passes, the number of babies named Allison decreases by 101.5. So, we can conclude that the name Allison has been decreasing in popularity over time (still a cool name though!).

# 5.8 Spelling by State

```{r}
baby_allan <- baby_names |>
  filter(Name %in% c("Allan", "Alan", "Allen"), Sex == "M")
```

```{r}
baby_allan |>
  group_by(Year, Name) |>
  summarise(Count = sum(Count), .groups = "keep") |>
  ggplot(mapping = aes(x = Year, 
                       y = Count, 
                       color = Name)) +
  geom_line() +
  labs(x = "Year Observed",
       y = " ",
       title = "Number of Male Born Babies Named Allan, Allen, or Alan")
```

# 5.9 Proportion of Allan, Allen, & Alan by State

```{r}
Ca_Pa <- baby_allan |>
  filter(Year == 2000, State %in% c("PA", "CA")) |>
  mutate(Prop = Count/sum(Count)) |>
   select(-Sex, -Year, -Count) |>
   pivot_wider(names_from = Name, 
              values_from = Prop,
              values_fill = 0)
Ca_Pa
 
```

#5.10 Percentages of Allan, Allen, & Alan by State

```{r}
Ca_Pa |>
  gt() |>
  fmt_percent(columns = 2:4, decimals = 0)
```
