---
title: "Lab 4: Millennials & Avocados"
author: "Daltzwer Oliveros"
format:
  html: 
    self-contained: true
    code-tools: true 
    code-fold: true 
execute:
  echo: true
  messages: false
  error: true
  eval: true
  warning: false
editor: visual
---

```{r Set up}
library(tidyverse)
library(here)
avocado <- read_csv(here("Week 4", "Labs", "avocado.csv")) |>
  filter(region != "TotalUS")
```

# Question 1.

This data contains information about the average price of avocados and the amount sold for each specific regions/cities.

# Question 2.

```{r Distinct region}

avocado |>
  distinct(region)
```

```{r rename avocado }
avocado <- avocado |>
  rename("Smalls" = `4046`, "Larges" = `4225`, "ExtraLarges" = `4770`)
```

```{r major region}
avocado_major <- avocado |>
  filter(region %in% c("Northeast", "Midsouth", "Southeast", "West", "SouthCentral", "Plains", "GreatLakes", "California")
                       )
```

```{r metro region}
avocado_metro <- avocado |>
  anti_join(avocado_major, by = "region")

```

# Question 3.

```{r}
avocado_major |>
  filter(type == "organic", year == "2017") |>
  select(Smalls, region) |>
  group_by(region) |>
  summarise(total_small_sold = sum(Smalls))
  

```

The West sold the most organic, small Hass avocados in 2017 at 1,870,206.3.

# Question 4.

```{r}
avocado_major |>
  separate(Date, c("Year", "Month", "Day"), 
           sep = "-") |>
  group_by(Month) |>
  summarise(avg_sales = mean(`Total Volume`)) |>
  slice_max(avg_sales)
```

On average, the month of February sees the highest volume of avocado sales at 2,600,098.

# Question 5

```{r}
top_five <- avocado_metro|>
  group_by(region) |>
  summarise(avg_sales = mean(`Total Volume`)) |>
  slice_max(avg_sales, n= 5) |>
  pull(region)
  

avocado_metro |>
  filter(region %in% top_five)|>
  ggplot(mapping = aes(x = region, y = `Total Volume`)) +
  geom_boxplot(fill = "lightblue", colour = "darkgreen") +
  labs(x = "Metro Regions", y = "Total Number of Avocados Sold")

```

As seen above, Los Angeles sold the most total avocados out of all the Metro Regions in our data set.

# Question 6

```{r California}
California <- avocado_metro |>
  filter(region %in% c("LosAngeles", "SanDiego", "Sacramento", "SanFrancisco")) |>
  group_by(region, type) |>
  summarise(AvocadoPrice = mean(AveragePrice), 
            .groups = "keep") |>
  pivot_wider(names_from = type,
              values_from = AvocadoPrice) |>
  mutate(PriceDifference = organic - conventional) |>
  arrange(desc(PriceDifference))
California

 
```

```{r}
California |>
  ggplot(mapping = aes(x = region, 
                       y = PriceDifference, 
                       fill = region)) +
  geom_col() +
  labs(x = "Regions of CA", 
       y = "Price Difference of Organic and Conventional")
```

According to the information above, San Francisco sees the biggest difference in the price of their organic avocados and conventional avocados at about \$0.81 followed by Sacramento at about \$0.70 difference.

# Question 7

```{r}
California |>
  inner_join(avocado_metro, by = "region") |>
  pivot_longer(cols = (Smalls:ExtraLarges),
               names_to = "Avocado Size", 
               values_to = "Amntsld") |>
  group_by(region, type, `Avocado Size`) |>
  summarise(avgsld = mean(Amntsld)) |>
  mutate(`Avocado Size` = factor(`Avocado Size`, 
                                 levels = c("Smalls", "Larges", "ExtraLarges"))) |>
  ggplot( aes(fill = `Avocado Size`, y = avgsld, x = region)) + 
    geom_bar(position = "fill", stat = "identity") +
  facet_wrap( ~ type) +
 scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  labs(x = "Region of CA", y = "Proportion of Mean Avocados Sold") +
  scale_fill_manual(values = c("lightblue", "dodgerblue2", "darkolivegreen2")) 

#https://r-graph-gallery.com/48-grouped-barplot-with-ggplot2
#https://datavizpyr.com/how-to-dodge-overlapping-text-on-x-axis-labels-in-ggplot2/
#http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf
  
```
